Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-06-15 doc/thesis/methods_results.Rnw
So much exposition I hate it

>>>>

\add{Regardless of how useful or interesting the \gls{BA} model parameters are, 
there is another aspect to the ``usefulness'' of our method - identifiability. 
Roughly speaking, the identifiability of a parameter says how much information
about that parameter can possibly be obtained from the observed data (see
\cref{chp:prelim}). If the parameters of the \gls{BA} model do not influence
tree shape at all, then the posterior distribution will exactly resemble the
prior. In that case, no matter how accurate a representation of the posterior
we are able to produce, no useful information can be learned about the network.
Hence, before proceeding with a full validation of \software{netabc} on
simulated data, we undertook two experiments designed to assess the
identifiability of each parameter separately.}

\add{First, we simulated trees under three different values of each parameter,
and asked how well we could tell the different trees apart. The better we are
able to distinguish the trees, the more identifiability we might expect for the
corresponding parameter when we attempt to estimate it with \gls{ABC}. In this
experiment, we fix all but one of the parameters at a time, so it is possible
that identifiability will be lost when all parameters are free to vary. This
experiment also has the secondary purpose of validating our choice of the tree
kernel as a distance measure in \gls{ABC}. To tell the trees apart, we use a
classifier based on the tree kernel, but we also test two other tree shape
statistics: Sackin's index, which considers only the topology, and the
\gls{nltt}, which considers only branching times. Since the tree kernel
incorporates both of these sources of information, we expect it to outperform 
the other two statistics. Finally, the tree kernel can be ``tuned'' by
adjusting the values of the meta-parameters \gls{lambda} and \gls{sigma}. We
used this experiment to choose values for these meta-parameters which were best
able to distinguish the different trees. }

\add{A second experiment was designed to test whether we could actually
estimate the parameters numerically, rather than just telling three different
values apart, and also to assess how identifiability varied in different
regions of the parameter space. An individual tree was compared to simulated
trees on a one-dimensional grid of values of one \gls{BA} parameter, to obtain
a ``distribution'' of kernel score values. From this distribution, estimates
and credible intervals of the parameter in question could be calculated. We
still held all but one parameter fixed, as the curse of dimensionality made it
infeasible to perform grid search on the full parameter space. Repeating this
experiment with trees located throughout the parameter space allowed us to
beter quantify the identifiability. Furthermore, doing marginal estimation
(that is, estimating one parameter with all others fixed) can provide insight
into any biases observed when doing joint estimation with \gls{ABC}. If grid
search is inaccurate, it indicates a lack of parameter identifiability.
However, if the marginal grid search estimates are accurate but the estimates
obtained with \gls{ABC} are biased, this points to confounding between the
parameters which could only be observed when they are all estimated jointly.}

\add{After these preliminary experiments, the strategy we used for testing
\software{netabc} was a standard simulation-based validation. Transmission
trees were simulated under several combinations of parameter values; we tried
to recover these values by running \software{netabc} on the simulated trees.}

\add{We have argued that the \gls{BA} parameters are worth estimating, and
described three simulation-based experiments to assess our ability to do so. 
There is a final aspect of the usefulness of these estimates: the accuracy of
the \gls{ABC} approximation to the posterior.}

... There are two factors which can affect the accuracy of the \gls{ABC}
approximation: the choice of distance function, and the settings used for
\gls{SMC} such as the number of particles and \gls{ESS} decay rate
\gls{alphaess}. Fortunately, these are both easy to test. If the \gls{ABC}
estimates are wrong, we can check if the distance function is responsible by
simulating several trees under both the estimated and true values. ...
For the \gls{SMC} parameters, we can simply rerun one simulation with better
settings (more particles, more samples per particle, and higher
\gls{alphaess}), and see if the approximation improves.
